---
title: "Stats Homework 7"
author: "Hanna Butsko"
date: "2023-10-27"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1

5.23
a) Possible outcomes: {YS, YD, NS, ND}
b) -i P(D) = total number of people that died divided by total count of people. 2111 / 577006 = 0.0037
P(D) = 0.4 %
```{r}
2111 / 577006
```

-ii P(N) = total number of people that were not wearing seat belt divided by total count of people. 164128 / 577006 = 0.2844
P(N) = 28 %
```{r}
164128 / 577006
```
c) For this we would have to find an intersection P(N & D) = 1601 / 577006 = 0.0028
P(N & D) = 0.3 %
```{r}
1601 / 577006
```
d) If N and D were independent, then P(N & D) = P(N) * P(D) = (164128 / 577006) * (2111 / 577006) = 0.0010
Since P(N & D) actually equaled 0.0028, N and D were not independent.
In context of this data it means that there is dependence between whether a person died in a car accident or no based on if they were wearing a seat belt or not.
```{r}
(164128 / 577006) * (2111 / 577006)
```

5.27
a) There are 8 possible outcomes: {YYY, YYN, YNY, YNN, NYY, NYN, NNY, NNN} where Y = sale and N = no sale.
Based on the information provided, 20 % of the people that enter the shop actually end up buying something. So probability of Y = 0.2 and probability of N = 0.8.
We can calculate probability of each outcome like so:
YYY = 0.2 * 0.2 * 0.2 = 0.008
YYN = 0.2 * 0.2 * 0.8 = 0.032
YNY = 0.2 * 0.8 * 0.2 = 0.032
YNN = 0.2 * 0.8 * 0.8 = 0.128
NYY = 0.8 * 0.2 * 0.2 = 0.032
NYN = 0.8 * 0.2 * 0.8 = 0.128
NNY = 0.8 * 0.8 * 0.2 = 0.128
NNN = 0.8 * 0.8 * 0.8 = 0.512
b) To find the probability of an event of at least one sale to the three customers, instead of listing all the outcomes that fit that category and sum up their probabilities, it will be easier to take the probability of an outcome with no sales and subtract it from 1, because in this case {NNN} will be a complement of an event of at least one sale to the three customers.
So, 1 - P({NNN}) = 1 - 0.512 = 0.488.
The probability of an event of at least one sale to the three customers equals to 0.488.
c) In part b, our probability calculations assumed that all the outcomes were independent of each other. However, in reality, the three customers who enter the shop might be friends or family members. In such a scenario, we can make the assumption that only one of them will make a purchase, essentially paying for the entire group. Nevertheless, this assumption can impact our probability calculation because, in technical terms, only one person out of the three actually made a purchase. Our initial calculations did not consider the possibility that each person might have acquired something from the store, even though only one person was responsible for the payment.

5.30
a) First, we would limit ourselves to the lowest income category space, disregarding other values that may distract us. Then, looking through the lowest income row, we observe two values: one representing probability of being audited, and the other one not. The question is asking to find the probability of being audited, so we would select the second value corresponding to "yes" which is 0.0085. Now, to obtain the conditional probability, we need to divide 0.0085 by the row marginal total probability, which would be 0.0085/(0.0085 + 0.9556) = 0.0088. Tax payer has 0.9 % chance to be audited given that they are in lowest income category.
```{r}
0.0085/(0.0085 + 0.9556)
```
b) Similar to the first question, we will limit ourselves to "yes" audited column. The value we are interested in is in lowest income category row, which is 0.0085. Now, we need to calculate row's marginal total probability, which would be 0.0085 + 0.0009 + 0.0003 = 0.0097. Now find the proportion: 0.0085/0.0097 = 0.8763. Given that the taxpayer was audited, they have 88 % chance to be in lowest income category.
```{r}
0.0085 / 0.0097
```

5.39
a) The probability that an adult is very happy in his or her relationship will be equal to total number of very happy adults divided by a total number of individuals that participated in the survey: 147 / 317 = 0.4637
P(VH) = 46 % where VH = very happy
```{r}
147 / 317
```
b) -i P(VH | M) = P(VH & M) / P(M) where VH = very happy and M = male.
P(VH & M) = 69 / 317
P(M) = 146 / 317
P(VH | M) = (69 / 317) / (146 / 317) = 0.4726
P(VH | M) = 47 %
```{r}
(69 / 317) / (146 / 317)
```
-ii P(VH | F) = P(VH & F) / P(F) where VH = very happy and F = female
P(VH & F) = 78 / 317
P(F) = 171 / 317
P(VH | F) = (78 / 317) / (171 / 317) = 0.4561
P(VH | F) = 45.6 %
```{r}
(78 / 317) / (171 / 317) 
```
c) Events VH and M are independent if P(VH | M) = P(VH)
P(VH | M) = 46 %
P(VH) = 46 %
P(VH | M) = P(VH) = 46 %, proving that these two events are independent.

5.40
a) To find the probability that the tennis player will serve good on the first try, we need to divide the total number of her good first serves by a number of the total serves. 28 / 41 = 0.6829
Probability is equal to 68 %.
```{r}
28 / 41
```
b) Because these two events are independent from each other, meaning that probability of faulting on the first serve does not influence the probability of faulting on the second serve, we can use the formula: P(A and B) = P(A) * P(B).
Find probability of first fault: we know that total amount of serves was 41 and amount of good first serves was 28, so amount of faulty first serves is 41 - 28 = 13. Now, probability of a faulty first serve would be: 13 / 41.
Find the probability of a double fault: we are given the information that she had three double faults during that game, so the probability would be 3 / 41.
Now lets apply the formula: (13 / 41) * (3 / 41) = 0.0232.
Probability of double faulting given that her first serve resulted in fault is 2.3 %
```{r}
(13 / 41) * (3 / 41)
```
c) 3 / 41 * 100 = 7 %
```{r}
3 * 100 / 41
```

5.48
a) The probability of at least two students in class of 25 people having the same birthday is 57 %, based on our calculations in class. The probability of two students having birthday on January 1st would be much lower than our previous example, because we are restricting the number of possible matches to a specific date. In the previous example, we assumed no specific birthday, meaning any student's birthday could match to any other student's birthday, which leads to a higher overall probability. However, if my birthday is on January 1st, that means that there are only 24 other students in class whose birthday could potentially match mine. This reduction in potential matches decreases the probability.
b) To make calculations easier, we can follow the same steps as we did when calculated birthday match probability in class. First, we should find the probability of no matches at all. If my birthday is on January 1st, then every other student that has birthday on any other day out of the year is guaranteed to not have a match with my birthday. So, any of 364 days out of 365 days will not match to my birthday. Because we are left with 24 students(25 - 1(me)), we need to multiply 364/365 24 times by itself: 364/365 * 364/365 * ... * 364/365, or (364/365)^24 = 0.9362772.
This is the probability of 24 out of 25 students with birthday that does not match mine. To find P(student with a birthday that matches mine) = 1 - 0.9362772 = 0.0637228 or approximately 6 %.
This proves that probability of finding a student with the same birthday as mine is much lower than finding any two students with the same birthday.
```{r}
(364/365)^24
```

5.50
a) Since we were given probabilities of hole in one for each hole, we can easily calculate complements for every hole in one:
P complement of hole in one on hole 4 = 1 - 0.0005 = 0.9995
P complement of hole in one on hole 6 = 1 - 0.0015 = 0.9985
P complement of hole in one on hole 12 = 1 - 0.0005 = 0.9995
P complement of hole in one on hole 16 = 1 - 0.0025 = 0.9975
We know that probabilities of making hole in one are independent from one hole to the nest, so, to find probability of no holes in one during the round of golf, we can just multiply the complements we calculated above: P(no hole in one) = 0.9995 * 0.9985 * 0.9995 * 0.9975 = 0.995008 or 99.5 %.
```{r}
0.9995 * 0.9985 * 0.9995 * 0.9975
```
b) Since every round is independent, to calculate probability for the next 20 rounds, we can use the result from part a) for every no hole in one round 20 times.
So the calculation would be P(no hole in one) * P(no hole in one) * ... * P(no hole in one) of a total 20 time, which is the same as P(no hole in one)^20.
P(no hole in one in 20 rounds) = P(no hole in one)^20 = 0.995008^20 = 0.9046105 or 90 %.
```{r}
0.995008^20
```
c) Because we already calculated the complement of making at least one hole in one during the next 20 rounds, it is easier to subtract that value from 1 to get the result, rather than recalculate the whole new probability.
P(at least one hole in one in 20 rounds) = 1 - 0.9046105 = 0.0953895 or 9.5 %.
```{r}
1 - 0.9046105 
```

5.57
a) S is yes for having breast cancer, S^C is a complement of S:
P(POS|S) = 0.86 * 0.01 = 0.0086 or 0.9 %
P(NEG|S) = (1 - 0.86) * 0.01 = 0.0014 or 0.1 %
P(POS|S^C) = (1 - 0.88) * (1 - 0.01) = 0.12 or 12 %
P(NEG|S^C) = 0.88 * (1 - 0.01) = 0.87 or 87 %
```{r}
0.86 * 0.01 
(1 - 0.86) * 0.01 
(1 - 0.88) * (1 - 0.01) 
0.88 * (1 - 0.01) 
```
b) To calculate P(POS) we have to apply formula: P(POS) = P(POS & S) + P(POS & S^C)
P(POS) = 0.0086 + 0.1188 = 0.1274 or 13 %.
c) The question is asking to find P(S|POS). To calculate this, we have to apply Bayers' rule: P(S|POS) = P(POS|S) * P(S) / P(POS) = 0.0086 * 0.01 / 0.1274 = 0.000675 or 0.07 %.
```{r}
0.0086 * 0.01 / 0.1274
```
d) Frequencies on branches of Breast Cancer are 1 for Yes and 99 for No. To calculate those values, we only need to know one of them, the other will equal to 100 minus the first one, because Yes and No values have to sum up to 100, which is the total number of women in the group. For diagnostic test, the sum of POS and NEG results has to equal to the total number of the corresponding branch, either Yes for breast cancer or No for no breast cancer. So, for example, POS result given no condition(12) plus NEG result given no condition(87) is equal to 99, which is the frequency of No breast cancer branch. 
P(S|POS) can be calculated using conditional probability formula, however, because we are already given all the frequencies, all we have to do is to plug in the numbers: 
POS = 1 + 12 = 13
S = 1
P(S|POS) = 1 / 13 = 0.08
The reason why the predicted probability of having breast cancer given that result of the test is positive is only 8%( which is quite low), is not because the test is not reliable, it caused by the fact that the prevalence is extremely low. We were told that only 1 out of 100 women in the group had breast cancer, which is 1 %. So, the probability of having the condition even though the test was positive is only 8 % because a very small percentage of women actually had cancer in the first place.

5.62
a) Probability that down syndrome occurs: 54 / 5282 = 0.01 or 1 %.
```{r}
54 / 5282
```
b) i. Sensitivity P(POS|S) = P(POS & YES) / P(YES) = (48/5282) / (54/5282) = 0.89 or 89 %
```{r}
(48/5282) / (54/5282)
```

ii. Specificity P(NEG|NO) = P(NEG & NO) / P(NO) = (3921/5282) / (5228/5282) = 0.75 or 75 %
```{r}
(3921/5282) / (5228/5282)
```
c) i. P(YES|POS) = P(YES & POS) / P(POS) = (48/5282) / (1355/5282) = 0.035 or 3.5 %
```{r}
(48/5282) / (1355/5282)
```
ii. P(NO|NEG) = P(NO & NEG) / P(NEG) = (3921/5282) / (3927/5282) = 0.998 or 99.8 %
```{r}
(3921/5282) / (3927/5282)
```
d) The probability of having a down syndrome and testing positive on a test is predicted to be 89 %, which is quite high. The probability of not having down syndrome given that test was negative is 99.8 %, which is also very high. And probability of not having the condition and testing negative is 75 %. By looking at these probabilities we can agree that the test is reliable in determining whether a child will have down syndrome. However, answer in part c) i. is telling us something different. It showed that estimated probability of actually having the condition given that test was positive is only 3.5 %, which is extremely low. This, however, does not indicate that test is ineffective, because if we look at the prevalence, which is P(YES) = 54/5282 = 0.01 or probability of child actually having down syndrome is only 1 %. This is very low, hence, the P(YES|POS) is also low, not because the test is not reliable, but because in reality a very small percentage of population actually has down syndrome.

6.5
a) P(1) = 0.10
P(2) = 0.30
P(3) = 0.40
P(4) = 0.20
b) To calculate the mean we need to multiply each possible value of the random variable by its probability and then add all these products.
1 * 0.1 + 2 * 0.3 + 3 * 0.4 + 4 * 0.2 = 2.7
An average final grade score assigned by instructor is 2.7.
```{r}
1 * 0.1 + 2 * 0.3 + 3 * 0.4 + 4 * 0.2
```

6.12
a) Let W represent winning and L represent losing. Sample space: {WW, WL, LW, LL}.
b) Probabilities:
First Item:
  For winning: 0.1
  For losing: 0.9
Second Item:
  For winning: 0.2
  For losing: 0.8
So, to calculate probability of outcomes in our sample space, we have to multiply corresponding probabilities:
P(WW) = 0.1 * 0.2 = 0.02
P(WL) = 0.1 * 0.8 = 0.08
P(LW) = 0.9 * 0.2 = 0.18
P(LL) = 0.9 * 0.8 = 0.72
c) For each outcome we calculate total amount of money spent: X = {50, 30, 20, 0}.
50 for WW, 30 for WL, 20 for LW, and 0 for LL.
P(50) = 0.02
P(30) = 0.08
P(20) = 0.18
P(0) = 0.72
d) Mean of X = 0.02 * 50 + 0.08 * 30 + 0.18 * 20 + 0.72 * 0 = 7 $
This represents the average amount of money we can expect to spend when participating in two eBay auctions with given probabilities of winning each bid.
```{r}
0.02 * 50 + 0.08 * 30 + 0.18 * 20 + 0.72 * 0
```
